### 2023-02-08 R_BIOLIB_slovniky.Rmd
BIOLIB animalia init

---
title: "MIGRACE DAT - DEMUS / MUSEION"
subtitle: "Prirodovedne sbirky - BIOLIB slovniky"
author: "Katerina Krizova"
date: "2023-02-08"
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, fig.dim = c(8, 4))

# PACKAGES

# install.packages("tidyverse", dependencies = T)
# install.packages("readxl")
# install.packages("stringr")
# install.packages("mdbr")
# install.packages("RODBC")
# install.packages("odbc")
require(tidyverse)
require(readxl)
require(stringr) # ::str_split_fixed
require(RODBC) # ::read_mdb
require(odbc) # ::read_mdb

```

\newpage

# PROJEKT A CESTY

```{r PROJECT SPECIFICATIONS AND PATHS}

# *** SWITCH *** SWITCH *** SWITCH *** SWITCH *** SWITCH ***

project <- "MSK_MUZBE"

# sbirka <- 
# podsbirka <-

# *** SWITCH *** SWITCH *** SWITCH *** SWITCH *** SWITCH ***

path <-  "C:/Users/krizova/Documents/R_migrace/"

path_data <-  paste0(path, "data/")
path_csv <-  paste0(path, "mus_csv/", project, "/")

path_mdb <-  "M:/02 dokumenty/03 migrace/"

```

# PRIPOJENI DATABAZE

```{r SET DATABASE CONNECTION}

dbname <- paste0(path_mdb, "BIO_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

RODBC::sqlTables(con)                                            # list tables in database

```

\newpage

# LOAD DB TABLES

## ANIMALIA

```{r READ AND MODIFY DB TABLES FOR ANIMALIA, warning = F}

# read all taxons

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen")
trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Animalia_Trida")  
rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Animalia_Rad")  
celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Animalia_Celed")  
rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Animalia_Rod")  
druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Animalia_Druh")  
poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Animalia_Poddruh")

```

## FUNGI

```{r READ AND MODIFY DB TABLE FUNGHI, warning = F}

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Fungi_Kmen")  
trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Fungi_Trida")  
rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Fungi_Rad")  
celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Fungi_Celed")  
rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Fungi_Rod")  
druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Fungi_Druh")  
poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Fungi_Poddruh")

```

## PLANTAE

```{r}

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Plantae_Kmen")  
trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Plantae_Trida")  
rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Plantae_Rad")  
celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Plantae_Celed")  
rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Plantae_Rod")  
druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Plantae_Druh")  
poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Plantae_Poddruh")

```

# CREATE SUMMARY TABLE

```{r KINGDOM-SPECIFIC}

# add taxon prefix to column names

colnames(kmen) <- paste0("kmen_", colnames(kmen))
colnames(trida) <- paste0("trida_", colnames(trida))
colnames(rad) <- paste0("rad_", colnames(rad))
colnames(celed) <- paste0("celed_", colnames(celed))
colnames(rod) <- paste0("rod_", colnames(rod))
colnames(druh) <- paste0("druh_", colnames(druh))
colnames(poddruh) <- paste0("poddruh_", colnames(poddruh))

# merge taxon tables one by one

ani1 <- merge(kmen, trida, by.x = "kmen_Nomen", by.y = "trida_NadNomen", all.y = T)
ani2 <- merge(ani1, rad, by.x = "trida_Nomen", by.y = "rad_NadNomen", all.y = T)
ani3 <- merge(ani2, celed, by.x = "rad_Nomen", by.y = "celed_NadNomen", all.y = T)
ani4 <- merge(ani3, rod, by.x = "celed_Nomen", by.y = "rod_NadNomen", all.y = T)
ani5 <- merge(ani4, druh, by.x = "rod_Nomen", by.y = "druh_NadNomen", all.y = T)
ani6 <- merge(ani5, poddruh, by.x = "druh_Nomen", by.y = "poddruh_NadNomen", all = T)

# remove prelim dfs

rm(ani1)
rm(ani2)
rm(ani3)
rm(ani4)
rm(ani5)
```