### 2023-02-13 _GEN_BIOLIB_slovniky.Rmd

---
title: "MIGRACE DAT - DEMUS / MUSEION"
subtitle: "Prirodovedne sbirky - BIOLIB slovniky"
author: "Katerina Krizova"
date: "2023-02-08"
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, fig.dim = c(8, 4))

# PACKAGES

# install.packages("tidyverse", dependencies = T)
# install.packages("readxl")
# install.packages("stringr")
# install.packages("mdbr")
# install.packages("RODBC")
# install.packages("odbc")
require(tidyverse)
require(readxl)
require(stringr) # ::str_split_fixed
require(RODBC) # ::read_mdb
require(odbc) # ::read_mdb

```

\newpage

# PROJEKT A CESTY

```{r PROJECT SPECIFICATIONS AND PATHS}

# *** SWITCH *** SWITCH *** SWITCH *** SWITCH *** SWITCH ***

# slovnik <- "animalia"
slovnik <- "fungi"
# slovnik <- "plantae"

# *** SWITCH *** SWITCH *** SWITCH *** SWITCH *** SWITCH ***

path <-  "C:/Users/krizova/Documents/R/"

path_mdb <-  "M:/02 dokumenty/03 migrace/BIO/"

```

# PRIPOJENI DATABAZE

```{r SET DATABASE CONNECTION}

dbname <- paste0(path_mdb, "bio_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

# RODBC::sqlTables(con)                           # list tables in database

```

\newpage

# LOAD DB TABLES

```{r READ AND MODIFY DB TABLES, warning = F}

if(slovnik == "animalia"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Animalia_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Animalia_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Animalia_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Animalia_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Animalia_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Animalia_Poddruh") 
} else if(slovnik == "fungi"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Fungi_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Fungi_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Fungi_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Fungi_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Fungi_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Fungi_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Fungi_Poddruh")
} else if(slovnik == "plantae"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Plantae_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Plantae_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Plantae_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Plantae_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Plantae_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Plantae_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Plantae_Poddruh")
} else print("I have a bad feeling about this!")

```

# CREATE SUMMARY TABLE

```{r KINGDOM-SPECIFIC}

# add taxon prefix to column names

colnames(kmen) <- paste0("kmen_", colnames(kmen))
colnames(trida) <- paste0("trida_", colnames(trida))
colnames(rad) <- paste0("rad_", colnames(rad))
colnames(celed) <- paste0("celed_", colnames(celed))
colnames(rod) <- paste0("rod_", colnames(rod))
colnames(druh) <- paste0("druh_", colnames(druh))
colnames(poddruh) <- paste0("poddruh_", colnames(poddruh))

# full_join - this setup gives a full table WITH ALL and ORDERED COLUMNS

join1 <- full_join(kmen, trida, by = c("kmen_Nomen" = "trida_NadNomen"), 
                   keep = T, multiple = "any") # 111
join2 <- full_join(join1, rad, by = c("trida_Nomen" = "rad_NadNomen"), 
                   keep = T, multiple = "any") # 605
join3 <- full_join(join2, celed, by = c("rad_Nomen" = "celed_NadNomen"), 
                   keep = T, multiple = "any") # 6 569
join4 <- full_join(join3, rod, by = c("celed_Nomen" = "rod_NadNomen"), 
                   keep = T, multiple = "any") # 23 970
join5 <- full_join(join4, druh,by = c("rod_Nomen" = "druh_NadNomen"), 
                   keep = T, multiple = "any") # 108 313
join6 <- full_join(join5, poddruh, by = c("druh_Nomen" = "poddruh_NadNomen"), 
                   keep = T, multiple = "any") # 128 848

write.table(join6, paste0(path, "BioLib_KKsummary_", slovnik, ".csv"), sep = ";", row.names = F)

```

# p i s k o v i s t e

```{r sandbox, eval = F}

# merge taxon tables one by one

# ani1 <- merge(kmen, trida, by.x = "kmen_Nomen", by.y = "trida_NadNomen", all.y = T) # 98
# ani2 <- merge(ani1, rad, by.x = "trida_Nomen", by.y = "rad_NadNomen", all.y = T) # 581
# ani3 <- merge(ani2, celed, by.x = "rad_Nomen", by.y = "celed_NadNomen", all.y = T) # 6 515
# ani4 <- merge(ani3, rod, by.x = "celed_Nomen", by.y = "rod_NadNomen", all.y = T) # 19 643
# ani5 <- merge(ani4, druh, by.x = "rod_Nomen", by.y = "druh_NadNomen", all.y = T) # 103 970
# ani6 <- merge(ani5, poddruh, by.x = "druh_Nomen", by.y = "poddruh_NadNomen", all = T) # 124 505

# inner_join - wrong path

# ani1 <- inner_join(kmen, trida, by = c("kmen_Nomen" = "trida_NadNomen")) # 98
# ani2 <- inner_join(ani1, rad, by = c("trida_Nomen" = "rad_NadNomen")) # 557
# ani3 <- inner_join(ani2, celed, by = c("rad_Nomen" = "celed_NadNomen")) # 6 153
# ani4 <- inner_join(ani3, rod, by = c("celed_Nomen" = "rod_NadNomen")) # 18 324
# ani5 <- inner_join(ani4, druh,by = c("rod_Nomen" = "druh_NadNomen")) # 92 791
# ani6 <- inner_join(ani5, poddruh, by = c("druh_Nomen" = "poddruh_NadNomen")) # 25 972

# right_join - this setup gives exactely the same result as 'merge', BUT WITH ORDERED COLUMNS

# ani11 <- right_join(kmen, trida, by = c("kmen_Nomen" = "trida_NadNomen"), keep = T) # 98
# ani22 <- right_join(ani1, rad, by = c("trida_Nomen" = "rad_NadNomen"), keep = T) # 557
# ani33 <- right_join(ani2, celed, by = c("rad_Nomen" = "celed_NadNomen"), keep = T) # 6 153
# ani44 <- right_join(ani3, rod, by = c("celed_Nomen" = "rod_NadNomen"), keep = T) # 18 324
# ani55 <- right_join(ani4, druh,by = c("rod_Nomen" = "druh_NadNomen"), keep = T) # 92 791
# ani66 <- left_join(ani5, poddruh, by = c("druh_Nomen" = "poddruh_NadNomen"), keep = T) # 25 972

check_kt <- join6 %>% 
  select(kmen_Nomen, trida_NadNomen, trida_Nomen) %>% 
  distinct()

check_tr <- join6 %>% 
  select(trida_Nomen, rad_NadNomen, rad_Nomen) %>% 
  distinct() %>% 
  filter(is.na(trida_Nomen))

check_rc <- join6 %>% 
  select(rad_Nomen, celed_NadNomen, celed_Nomen) %>% 
  distinct() %>% 
  filter(is.na(rad_Nomen))

write.table(check_rc, paste0(path_csv, "check_tabs2print_rc.csv"), sep = ";", row.names = F)

check_tab <- join6 %>% 
  select(kmen_Nomen, trida_NadNomen, trida_Nomen, 
         rad_NadNomen, rad_Nomen, 
         celed_NadNomen, celed_Nomen, 
         rod_NadNomen, rod_Nomen, 
         druh_NadNomen, druh_Nomen, 
         poddruh_NadNomen, poddruh_Nomen, 
         )

check_tab2 <- check_tab %>% 
  select(-poddruh_Nomen, -poddruh_NadNomen, -druh_Nomen) %>% 
  filter(is.na(kmen_Nomen)) %>% 
  distinct()

# write.table(check_tab2, paste0(path_csv, "check_kmeny.csv"), sep = ";")

check_kmen <- check_tab %>% 
  select(kmen_Nomen, trida_NadNomen) %>% 
  distinct() %>% 
  filter(is.na(trida_NadNomen))

check_trida <- check_tab %>% 
  select(trida_Nomen, rad_NadNomen) %>% 
  distinct() 
  filter(is.na(rad_NadNomen))

check_chordata <- join6 %>% 
  filter(rad_NadNomen %in% c("Chordata"))

```