### 2023-03-01 _GEN_axi_piskoviste.Rmd
MDH finalized

---
title: "AXIELL piskoviste"
subtitle: "Prostredi pro zkouseni novych postupu"
author: "Kateřina Křížová"
date: "2023-01-26"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

```{r}
require(tidyverse)
```

\newpage

# *GENERAL*

## X M L

### import

PACKAGES: "XML", "xml2"  
ARTICLES: https://appsilon.com/r-xml/  ;  https://www.w3schools.com/xml/schema_howto.asp

```{r XML IMPORT}

install.packages("XML")
require(XML)

install.packages("xml2")
require(xml2)

install.packages("methods")
require(methods)

install.packages("tibble")
require(tibble)

# funkcni prikazy
xml <- xml2::read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/grafika.xml", as_html = F) 
xsd <- xml2::read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/grafika_xsd.xsd")

xml2::xml_validate(xml, xsd) # TRUE - what does this mean?
structure <- xml2::xml_structure(xml) # ?
namesp <- xml2::xml_set_namespace(xml, uri = "http://www.w3.org/XML/1998/namespace") #

xml <- xml2::xmlParse(xml)

dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/cvdata.xml") # ok
dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/grafikaupr.xml") # jiny format, nenacita spravne
dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/PRM_BioLib_090_Fungi_Trida.xml") # ok
print(dataframe)

# lokalni piskoviste

data <- XML::xmlParse(xml)
rootnode <- XML::xmlRoot(data)
nodes <- XML::xmlSize(rootnode)
second_node <-rootnode[2]
attri <- rootnode[[3]][[4]]

cat('number of nodes: ', nodes)
print ('details of 2 record: ')
print (second_node)
print ('3rd attribute of 4th record: ', attr)

# nefunkcni

xml_structure(xml) # doesnt work
xml_find_all(xml, ".//position") # doesnt work

dept <- XML::xml_text(xml_find_all(xml, ".//department"))
salary <- XML::xml_text(xml_find_all(xml, ".//salary"))

xmltib <- tibble(department = dept, salary = salary)

df_employees <- XML::xmlToDataFrame(nodes = XML::getNodeSet(xml, "//employee"))
```