### 2023-03-13 _GEN_axi_zeroScript.Rmd

---
title: "ZERO SCRIPT"
subtitle: "Sablona pro proces migrace dat"
author: "Kateřina Křížová"
date: "2023-01-26"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

```{r}

# L O A D

demus_tab <- read_excel()

# M O D I F Y

modif_tab <- demus_tab %>% 
  mutate(MEMPTY = "")

modif_save <- paste0(path_csv, spravce, "")

if(file.exists(modif_save)){
  print("You've been here before!")
} else {
  write.table(modif_tab, file = modif_save, 
            sep = ";", quote = T, row.names = F, 
            na = "", fileEncoding="cp1250")
  print("Operational file  S A V E D  !")
  }

# C R E A T E    C S V

mus_tab <- modif_tab

oper_technika <- mus_tab

# S A V E   I M P O R T   C S V

save <- paste0(path_csv, spravce, "")

if(file.exists(save)){
  print("You've been here before!")
} else {
  write.table(mus_tab, file = save, 
            sep = ";", quote = T, row.names = F, 
            na = "", fileEncoding="cp1250")
  print("Import csv  W R I T T E N  !")
  }

```