### 2023-04-19 _GEN_axi_Rhacks.Rmd

---
title: "AXIELL R HACKS"
subtitle: "Uzitecne prikazy a funkce pro migrace dat"
author: "Kateřina Křížová"
date: "2023-01-26"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

# *GENERAL*

## N U M E R I C S

### Leading zeros -add

```{r LEADING ZEROS}

# package
require(stringr) # ::str_pad

# command
mutate(MPORC = str_pad(PorC_S, 6, pad = "0"))

```

mutate = create a new column  
MPORC = name of a new column
str_pad = [function] (https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_pad)
PorC_S = name of a source column (contains value '6' which we want add lead.zeros to)
6 = we want a final value to have exactly 6 places  
pad = "0" = empty places fill with "0"

### Leading zeros -remove

```{r}

mutate(Mcislo = sub("^0+", "", Mcislo))
```

## T E X T

### String length

```{r}

mutate(Mpskp_len = nchar(pskup))

```

### String detect case INsensitive

```{r}

str_detect(MFN, regex("^FM\\s\\d+\\.jpg$", ignore_case = T))

```

### Split string

```{r SPLIT STRING}

# # Split name column into firstname and last name
# df[c('First Name', 'Last Name')] <- str_split_fixed(df$Name, ' ', 2)
#  
# # Rearrange columns and remove original name column
# df <- df[c('First Name', 'Last Name', 'State')]

separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ")

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge") # at first whitespace

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge", remove = F) # ponecha puvodni sloupec

```

### Keep text before specific sign

```{r}
  mutate(blMCELED = sub("\\/.*", "", fiMTREE)) # ponechej text do prvni /
```

### Extract specific word in a row

```{r}
mutate(boMROD = word(druh, 1)) # vezmi 1. slovo ze sloupce druh
```

### Remove parentheses

```{r}

mutate(celed_LAT = gsub("[()]", "", celed_LAT))

```

### Remove whitespace

```{r RMV WHITE SPACE}

# package
require(stringr) # ::str_pad

# command
mutate(MMATERIAL = str_trim(MMAT_AGG)) 
```

str_trim = [function](https://stringr.tidyverse.org/reference/str_trim.html)

### Create a string of multiple column values with NAs

```{r}

unite(sys_tree_BIOL, c("kmen", "trida", "rad", "celed", "rod", "druh_Nomen"),
      sep = "/", na.rm = T, remove = FALSE) 
```

unite = [function](https://www.rdocumentation.org/packages/tidyr/versions/0.8.2/topics/unite)  
sys_tree_BIOL = name of a new column  
c(***) = list of columns  
sep = sign for separation  
na.rm = skip columns containing NA  
remove = keep or delete source columns

\newpage

## T A B L E S

### Add taxon prefix to column names

```{r}
colnames(kmen) <- paste0("kmen_", colnames(kmen))
```

### Deal with duplicates

```{r}

#  https://community.rstudio.com/t/filter-out-all-rows-with-duplicate-values/41043/4

# Return names which have only a single row of data
data %>% 
  group_by(name) %>% 
  filter(n()==1)

# Return names which have more than one row of data
data %>% 
  group_by(name) %>% 
  filter(n() != 1)
```

### Merge two collums with not complete information

column 1 has some NAs  
column 2 has also some NAs but in different rows than column 1
TASK: create one column with merged info

```{r}
#package
require(dplyr) # part of tidyverse

# command
mutate(kmen = coalesce(kmen_Nomen, trida_NadNomen))

```

coalesce = [function](https://www.rdocumentation.org/packages/dplyr/versions/1.0.10/topics/coalesce)

### Relocate columns

```{r}
relocate(MINVC, .before = JineC_S) 
relocate(MINVCDO, .after = MINVC)
```

\newpage

## D A T A B A S E S

### Connect db and read table

```{r}
# package 
require(RODBC)

# commands
dbname <- paste0(path_mdb, "bio_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

RODBC::sqlTables(con)                          # list tables in database

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen") # read specified table

```

*.mdb = old MS ACCESS  
*.accdb = new MS ACCESS

### List files from one or more folders

```{r}

# one
vypis <- as.data.frame(list.files("D:/MUZBE/Negativy/Image", full.names = T))

# more
vypis <- as.data.frame(list.files(c("D:/MUZBE/Obecné sbírky-4složky obrázků/Image_01",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_02",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_03",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_04"), 
                                  pattern = "jpg$|JPG$", full.names = T))

colnames(vypis)[1] <- "path"

```

## J O I N S

*inner_join* only keeps observations from x that have a matching key in y
*left_join()* keeps all observations in x.
*right_join()* keeps all observations in y.
*full_join()* keeps all observations in x and y.

[description and parameters](https://dplyr.tidyverse.org/reference/mutate-joins.html)

\newpage

# *REAL TASKS*

## D A T A C E / upravit

```{r DERIVE VZNIKPTU FROM SBIRKY, warning = F, eval = F}

dopr_tab <- sbirky %>% 
  select(MINVC, Autor_S, Role_S, Datace_S, MistoVz_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate(MDATACE = Datace_S) %>% 
  # yyyy
  mutate(MDAT_Y = ifelse(str_detect(MDATACE, "_00_00"), 
                         str_replace_all(MDATACE, "_00|-00", ""), NA)) %>% 
  # mm.yyyy
  # mutate(MDAT_YM = ifelse(str_detect(MDATACE,
  #                                    "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
  #                                   # format(anydate(substring(MDATACE, 1, 7), "%m.%Y")), ""))
  #                                   # substring(format(anydate(MDATACE), "%m.%Y"), 1, 7), ""))
  #                                   format(anydate(MDATACE), "%m.%Y"), ""))
  # mutate(MDAT_YM = format(anydate(MDATACE), "%m.%Y")) %>%
  mutate(MDAT_YM = substring(ifelse(str_detect(MDATACE,
                                             "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
                                  MDATACE, ""), 1, 7)) %>%
  mutate(MDAT_YM = format(anydate(MDAT_YM), "%m.%Y")) %>%
  # dd.mm.yyyy
  mutate(MDAT_YMD = ifelse(str_detect(MDATACE,
                          "^([0-9]{4})[_]([0-9]{2})[_]([0-9]{2})$"), 
                           format(anydate(MDATACE), "%d.%m.%Y"), NA)) %>% 
  # yy leta ss stoleti
  mutate(MDAT_LETA = ifelse(str_detect(MDATACE, "^([0-9]{4})[_]XX[_]([0-9]{2})$"),
                                    paste0(substring(dopr_tab$MDATACE, 9, 9), "0. léta ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # poloviny
  mutate(MDAT_2 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d2[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". polovina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # tretiny - NEVYSKYTUJE SE
  mutate(MDAT_3 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # poloviny - NEVYSKYTUJE SE
  mutate(MDAT_4 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". čtvrtina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # tretina roku
  mutate(MDAT_3r = ifelse(str_detect(MDATACE, "^\\d([1-9]{2})\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina roku ",
                                           substring(dopr_tab$MDATACE, 1, 4)), NA)) %>% 
  # DATACE FINAL
  mutate(MDATACE_FIN = coalesce(MDAT_Y, MDAT_YM, MDAT_YMD, MDAT_LETA, MDAT_2, MDAT_3r)) %>% 
  mutate(MDATACE_FIN = coalesce(MDATACE_FIN, MDATACE))

```

## TVORBA INVENTARNIHO CISLA

```{r MINVC}

sbirky <- read_excel(paste0(path_data, "/Sbirky.xlsx")) %>%

  select(-MSUBOD, -MSUBDO, -MINVC, -MINVCDO) %>%    # get rid of .accdb columns

  mutate(MPORC = str_pad(PorC_S, 6, pad = "0"),
         MPORCDO = str_pad(HPorC_S, 6, pad = "0"),
         MSUB = ifelse(Lomeni_S == "_", NA, Lomeni_S),
         MSUBDO = HLomeni_S, 
         MINVC = case_when(is.na(MSUB) ~ paste0(Rada_S, "-", MPORC),
                             TRUE ~ paste0(Rada_S, "-", MPORC, "/", MSUB)),
         MINVCDO = case_when(is.na(MSUBDO) ~ "",
                             TRUE ~ paste0(Rada_S, "-", MPORCDO, "/", MSUBDO))) %>% 
  # select(-MPORC, -MPORCDO) %>%
  relocate(MINVC, .before = JineC_S) %>% 
  relocate(MINVCDO, .after = MINVC)

```

\newpage

# *NEED TO COVER*

dopr_tab <- sbirky %>% 
  select(MINVC, Technika_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate('2sbirkaCisloEvidInt' = sbirka) %>% 
  mutate('3podsbirkaCislo' = podsbirka) %>% 
  mutate(MTECH = Technika_S) %>% 
  separate(MTECH,into=c("MTECH1", "MTECH2", "MTECH3", "MTECH4", "MTECH5"),
           sep=",(?! tuší)") %>% # separator is ',' BUT ONLY IF NOT FOLLOWED BY ' tuší' -> solves a single case of 'kresba tužkou, tuší a bělobou' which should remain as one entry
  pivot_longer(
    cols = MTECH1:MTECH5, 
    names_to = "MTECH_POR", # poradi techniky
    values_to = "MTECH_AGG",
    values_drop_na = TRUE) %>% 
  mutate(MTECHNIKA = str_trim(MTECH_AGG)) %>% # remove white space
  select(-MTECH_AGG) %>% 
  mutate(MTECH_POR = substr(MTECH_POR, 6, 6))

//

plant <- db_tab %>%

   separate(celed,into=c("celed_LAT", "celed_CZ") , sep=" ") %>% 
   separate(druh,into=c("rod_LAT", "druh_LAT", "rest"), sep = "\\s(?!x )", 
            extra = "merge", remove = FALSE) %>% 
   mutate(druh_LAT = paste(rod_LAT, druh_LAT),
          celed_CZ = gsub("[()]", "", celed_CZ),
          autor = ifelse(str_detect(rest, "^[[:upper:]]") | str_detect(rest, "^\\W"), rest,""),
          druh_LAT = gsub("x$", "", druh_LAT))

//

 mutate_all( ~ str_replace_all(., pattern = ";", replacement = " ")) %>% # removes semicolons
  mutate_all( ~ str_replace_all(., pattern = "[\r\n]", replacement = " ")) #removes paragraphs