### 2023-04-26 DEMZ_xml_agenda.Rmd
MUZBE bot biokarta almost ready

---
title: "DEMZ"
subtitle: "Zpracovani xml souboru"
author: "Kateřina Křížová"
date: "2023-01-26"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

# D E M Z

```{r PACKAGES}

require(tidyverse)
require(XML)
require(xml2)
require(methods)

```

```{r PATHS}

path_sndb <- "C:/Users/krizova/Documents/R/data/_piskoviste/"
path_xml <- "P:/w4dz/migrace/"
path_xml2 <- "P:/w4dz/migrace/new-xml/e4g/"

```

## piskoviste

```{r PISKOVISTE}

# xml1 <- as_list(read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml"))
# xml2 <- as_list(read_xml("P:/w4dz/migrace/boj/26272175.xml"))
# xml3 <- as_list(read_xml("P:/w4dz/migrace/boj/26280130.xml"))
# xml4 <- as_list(read_xml("P:/w4dz/migrace/boj/26282696.xml"))
# xml5 <- as_list(read_xml("P:/w4dz/migrace/boj/26332340.xml"))

path_xml <- "P:/w4dz/migrace/"
path_xml2 <- "P:/w4dz/migrace/new-xml/e4g/"
# fold <- "boj"
# fold <- "zmt"
fold <- "stm"
Lx <- c("10918171.xml", "26272175.xml", "26280130.xml", "26282696.xml", "26332340.xml")

infoex <- data.frame()

for (i in Lx) {
  a <- as_list(read_xml(paste0(path_xml, fold, "/", i)))
  b <- tibble::as_tibble(a) %>%
  unnest_longer(import)
  c <- b %>%
  dplyr::filter(import_id == "PrirCisl") %>%
  unnest_wider(import, names_sep = "_")
  s <- c[1]
  infoex <- rbind(infoex, s)
}

# ---------- novy format 20042023 ----------

xml1 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml"))
xml2 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16809748.xml"))
xml3 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16810414.xml"))
xml4 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16820407.xml"))
xml5 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16906385.xml"))

xml1 <- XML::xmlParse("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
print(xml1)

rootnode <- XML::xmlRoot(xml1)
nodes <- XML::xmlSize(rootnode)
second_node <- rootnode[2]
print(second_node)
third_node <- rootnode[3]
print(third_node)
fourth_node <- rootnode[4]
print(fourth_node)

dfx <- XML::xmlToDataFrame(nodes = XML::getNodeSet(xml1, "//property")) # vraci prazdny df

xml1 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml"))
xml1_tb <- tibble::as_tibble(xml1) %>% 
  unnest_longer('mets')
xml1_df <- xml1_tb %>% unnest_wider("mets") # nefunguje

xml1 <-read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
xml1_path <- xml_find_all(xml1, xpath = "//mets/metsHdr")
driver_id <- xml_attr(driver, "driverId")
url_name <- xml_attr(driver, "url")

#------------ tuto ------------- https://appsilon.com/r-xml/

# R XML Basics – How to Read and Parse XML Files

library(xml2)
library(XML)

#load
employee_data <- read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/data.xml")
demz <- read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/priloha_C_1_priklad_1.xml")
employee_data
demz

# parse
employee_xml <- xmlParse(employee_data)
demz_xml <- xmlParse(demz)
employee_xml
demz_xml

xml_structure(employee_data) # Pro tip: if you don’t care about the data, you can print the structure only. That’s done with the xml_structure() function:
xml_structure(demz)

xml_find_all(demz, ".//e4g:name")
xml_find_all(xml1, ".//objectDefName") # nic

xml_text(xml_find_all(employee_data, ".//position")) # only values
xml_text(xml_find_all(demz, ".//e4g:Meno"))
xml_text(xml_find_all(demz, ".//e4g:Meno//eg4:name")) # nefunguje

# How to Convert XML Data to tibble and data.frame

library(tibble)

# Extract department and salary info
dept <- xml_text(xml_find_all(employee_data, ".//department"))
salary <- xml_integer(xml_find_all(employee_data, ".//salary"))

# Format as a tibble
df_dept_salary <- tibble(department = dept, salary = salary)
df_dept_salary

df_employees <- xmlToDataFrame(nodes = getNodeSet(employee_xml, "//employee"))
df_employees

# a <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml"))
a <- as_list(read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/priloha_C_1_priklad_1.xml"))
aa <- a$`OAI-PMH`$ListRecords
aaa <- aa$record$header
df <- as.data.frame(do.call(rbind, aa))
df2 <- unlist(df)

b <- tibble::as_tibble(aa)
  unnest_longer('OAI-PMH')
c <- b %>%
  dplyr::filter('OAI-PMH_id' == "record") %>%
  unnest_wider('OAI-PMH', names_sep = "_")
d <- unlist(c)
s <- c[1]
infoex <- rbind(infoex, s)

infoex <- data.frame() # info export dataframe

for (i in Lxml_stm) {
  a <- as_list(read_xml(paste0(path_xml2, fold, "/", i))) # read xml
  b <- tibble::as_tibble(a) %>%  
  unnest_longer(mets)
  c <- b %>%
  dplyr::filter(mets_id == "agent") %>%
  unnest_wider(mets, names_sep = "_")
  d <- unlist(c)
  s <- d[1]
  infoex <- rbind(infoex, s)
}

xml <- xml2::read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
xsd <- xml2::read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/priloha_c_1_CAIR_CEMUZ_e4g.xsd")

xml2::xml_validate(xml, xsd) # TRUE - what does this mean?

# ---------- Exploring and Extracting Data from XML in R ---------- 
# https://rstudio-pubs-static.s3.amazonaws.com/499292_d6edbb19b08f456097333fbf9443f9b7.html

xml3 <- XML::xmlParse("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
class(xml3)
xmlRoot(xml3)

xmltop <- xmlRoot(xml3) #gives content of root
xmlName(xmltop)
xmlSize(xmltop) # how many children in node,
xmlName(xmltop[[1]]) # name of root's children
xmlName(xmltop[[2]]) # name of root's children
xmltop[[1]]
xmltop[[2]]
xmltop[[3]]
xmltop[[4]]

xmlName(xmltop[[1]][[1]])
xmlName(xmltop[[1]][[1]][[1]])
xmlName(xmltop[[1]][[1]][[1]][[1]])
xmlValue(xmltop[[1]][[1]][[1]][[1]])
xmlValue(xmltop[[2]][[1]][[1]][[1]])

xpathSApply(xml3, '/PubmedArticleSet/PubmedArticle/MedlineCitation/PMID', xmlValue)
xpathSApply(xmltop[[2]], '//display', xmlValue)
xpathSApply(xml3, '/mets/dmdSec/mdwrap/xmlData/esez/objectInfo/display/', xmlValue)

#-------------------------- priklad-------------------------------

xml <- XML::xmlParse("C:/Users/krizova/Documents/R/data/_piskoviste/priloha_C_1_priklad_1.xml")

class(xml)
xmlRoot(xml)

xmltop <- xmlRoot(xml) #gives content of root
xmlName(xmltop)
xmlSize(xmltop) # how many children in node,
xmlName(xmltop[[1]]) # name of root's children
xmlName(xmltop[[2]]) # name of root's children
xmlName(xmltop[[3]]) # name of root's children
xmltop[[1]]
xmltop[[2]]
xmltop[[3]]

xmlName(xmltop[[1]][[1]])

xpathSApply(xml3, '/PubmedArticleSet/PubmedArticle/MedlineCitation/PMID', xmlValue)
xpathSApply(xmltop, '//e4g:name', xmlValue)
k <- xpathSApply(xml, '/OAI-PMH/ListRecords/record/metadata/e4g:E4GEvid2/e4g:Jazyk/e4g:id', xmlValue)
print(k)

```

```{r PISKOVISTE 2}

# read

xmla <- read_xml(paste0(path_sndb, "priloha_C_1_priklad_1.xml"))
xmlb <- xml2::as_list(xml2::read_xml(paste0(path_sndb, "priloha_C_1_priklad_1.xml")))
xmlc <- XML::xmlParse(paste0(path_sndb, "priloha_C_1_priklad_1.xml"))

# extract third element of nested list

# install.packages("purrr")            # Install & load purrr package
library("purrr")
a <- map(xmlb, 3)                      # https://statisticsglobe.com/select-first-element-of-nested-list-in-r
b <- map(a, 1)

c <- b[[1]][[2]]
c[[1]][[9]]

# realny xml, ne priklad

l <- xml2::as_list(xml2::read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml"))
x <- xml2::read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
xp <- XML::xmlParse(xml2::read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml"))

# xml2 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16809748.xml"))
# xml3 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16810414.xml"))
# xml4 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16820407.xml"))
# xml5 <- as_list(read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16906385.xml"))

library(XML)
myData <- xmlParse(xmla)
xmltop <- xmlRoot(myData)
xmlName(xmltop)
xmlSize(xmltop)

md <- xpathSApply(xp, "//*//e4gEvidCislo", # xpath works only for parsed files
             # , namespaces = c(e4g = "http://www.cemuz.sk/ver3/")  #*/e4g:E4GEvid2/
              xmlValue)[1]

md <- xpathSApply(xp, "mets/dmdSec/mdWrap/xmlData/esez/objectInfo/properties/e4gEvidCislo", # xpath works only for parsed files
             # , namespaces = c(e4g = "http://www.cemuz.sk/ver3/")  #*/e4g:E4GEvid2/
              xmlValue)[1]

md <- xpathSApply(myData, "//*/xs:EvidCislo"
              , namespaces = c(xs = "http://www.w3.org/2001/XMLSchema")
              , xmlValue)

xpathSApply(myData, "//*/e4g:E4GEvid2/e4g:EvidCislo"
              , namespaces = c(e4g = "http://www.cemuz.sk/ver3/")  #*/e4g:E4GEvid2/
              , xmlValue)[1]           # https://www.opencodez.com/how-to-guide/how-to-use-xpath-for-web-scraping-with-r.htm

xpathSApply(myData, "display"
             # , namespaces = c(e4g = "http://www.cemuz.sk/ver3/")  #*/e4g:E4GEvid2/
              , xmlValue)           # https://www.opencodez.com/how-to-guide/how-to-use-xpath-for-web-scraping-with-r.htm

# nic nefunguje, nechapu, nerozumim

```

```{r PRIKLAD}

xa <- read_xml(paste0(path_sndb, "priloha_C_1_priklad_1.xml"))
# xa <- read_xml("P:/w4dz/migrace/new-xml/e4g/stm/16788817.xml")
xb <- XML::xmlParse(paste0(path_sndb, "priloha_C_1_priklad_1.xml"))
# xl <- xml2::as_list(xml2::read_xml(paste0(path_sndb, "priloha_C_1_priklad_1.xml")))

# https://appsilon.com/r-xml/
# https://rpubs.com/Howetowork/499292

class(xa) # "xml_document" "xml_node"    
class(xb) # "XMLInternalDocument" "XMLAbstractDocument"

xmlRoot(xb)

# struct <- xml_structure(xa) # ne

xml_find_all(xa, ".//e4g:Muzeum/e4g:name") # !!!! FUNKCNI !!!!
xml_find_all(xa, ".//e4g:Muzeum/e4g:id") # !!!! FUNKCNI !!!!
xml_find_all(xa, ".//e4g:Jazyk/e4g:name") # !!!! FUNKCNI !!!!

muzec <- xml_text(xml_find_all(xa, ".//e4g:EvidCislo")) # !!!! FUNKCNI !!!!
muzname <- xml_text(xml_find_all(xa, ".//e4g:Muzeum/e4g:name")) # !!!! FUNKCNI !!!!
muzid <- xml_text(xml_find_all(xa, ".//e4g:Muzeum/e4g:id")) # !!!! FUNKCNI !!!!
muzos <- xml_text(xml_find_all(xa, ".//e4g:NadobOsoby/e4g:Rola/e4g:name")) # 1:n, ostatni maji 96 zaznamu, zde jiz 279 -> vyresit! 
# https://stackoverflow.com/questions/16805050/extract-xml-node-and-attribute-in-r

muz <- tibble(evidc = muzec, name = muzname, id = muzid) # !!!! FUNKCNI !!!!

```

## ostra analyza

*!!! NIZE JEN OVERENE A FUNKCNI PRIKAZY !!!*

```{r LOAD AND EDIT XML}

# L I S T   O F   F I L E S ^

Lxml_boj <- list.files("P:/w4dz/migrace/boj/", pattern = "xml") # 182
Lxml_zmt <- list.files("P:/w4dz/migrace/zmt/", pattern = "xml") # 59795
Lxml_stm <- list.files("P:/w4dz/migrace/new-xml/e4g/stm/", pattern = "xml") # 955

Lxml <- c(Lxml_boj, Lxml_zmt) # 59977

# fold <- "boj"
# fold <- "zmt"
fold <- "stm"

# L O O P   F O R   E X T R A C T I N G   S P E C.   I N F O

infoex <- data.frame() # info export dataframe

for (i in Lxml_boj) {
  a <- as_list(read_xml(paste0(path_xml, fold, "/", i))) # read xml
  b <- tibble::as_tibble(a) %>%  
  unnest_longer(import)
  c <- b %>%
  dplyr::filter(import_id == "Signatura") %>%
  unnest_wider(import, names_sep = "_")
  s <- c[1]
  infoex <- rbind(infoex, s)
}

```