### 2023-04-26 _GEN_axi_piskoviste.Rmd

---
title: "AXIELL piskoviste"
subtitle: "Prostredi pro zkouseni novych postupu"
author: "Kateřina Křížová"
date: "2023-01-26"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

```{r}
require(tidyverse)
```

\newpage

# *GENERAL*

## X M L

### import

PACKAGES: "XML", "xml2", "methods", "tibble"
ARTICLES: https://appsilon.com/r-xml/  ;  https://www.w3schools.com/xml/schema_howto.asp

```{r XML IMPORT}

install.packages("XML")
require(XML)

install.packages("xml2")
require(xml2)

install.packages("methods")
require(methods)

install.packages("tibble")
require(tibble)

# funkcni prikazy
xml <- xml2::read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml", as_html = F) 
xsd <- xml2::read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/priloha_c_1_CAIR_CEMUZ_e4g.xsd")

xml2::xml_validate(xml1, xsd) # TRUE - what does this mean?
structure <- xml2::xml_structure(xml) # ?
namesp <- xml2::xml_set_namespace(xml, uri = "http://www.w3.org/2001/XMLSchema-instance") #

xml <- xml2::xmlParse(xml)

dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml") # ok
dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/grafikaupr.xml") # jiny format, nenacita spravne
dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/PRM_BioLib_090_Fungi_Trida.xml") # ok
print(dataframe)

# lokalni piskoviste

data <- XML::xmlParse(xml)
rootnode <- XML::xmlRoot(data)
nodes <- XML::xmlSize(rootnode)
second_node <-rootnode[2]
attri <- rootnode[[3]][[4]]

cat('number of nodes: ', nodes)
print ('details of 2 record: ')
print (second_node)
print ('3rd attribute of 4th record: ', attr)

# nefunkcni

xml_structure(xml) # doesnt work
xml_find_all(xml, ".//position") # doesnt work

dept <- XML::xml_text(xml_find_all(xml, ".//department"))
salary <- XML::xml_text(xml_find_all(xml, ".//salary"))

xmltib <- tibble(department = dept, salary = salary)

df_employees <- XML::xmlToDataFrame(nodes = XML::getNodeSet(xml, "//employee"))

#----------------------------- B R N O -------------------------------------------

xml <- XML::xmlParse("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml") 
print(xml)

# rootnode <- XML::xmlRoot(xml)
# nodes <- XML::xmlSize(rootnode)
# second_node <- rootnode[2]
# attri <- rootnode[[4]][[3]]
# 
# dataframe <- XML::xmlToDataFrame("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml") # nefunguje

# now
# https://urbandatapalette.com/post/2021-03-xml-dataframe-r/

install.packages("xml2")
require(xml2)
restaurant_license_xml <- as_list(read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml"))

xml_df <- tibble::as_tibble(restaurant_license_xml) %>%
  unnest_longer(import)

lp_wider <- xml_df %>%
  dplyr::filter(import_id == "Podskup") %>%
  unnest_wider(import, names_sep = "_" )

lp_df = lp_wider %>%
  # 1st time unnest to release the 2-dimension list?
  unnest(cols = names(.)) %>%
  # 2nd time to nest the single list in each cell?
  unnest(cols = names(.)) %>%
  # convert data type
  readr::type_convert()

myFun <- function(data) {
  temp1 <- sapply(data, is.list)
  temp2 <- do.call(
    cbind, lapply(data[temp1], function(x) 
      data.frame(do.call(rbind, x), check.names=FALSE)))
  cbind(data[!temp1], temp2)
}

xml_unlist <- myFun(xml_df)
xml_unlist2 <- tidyr::unnest_longer(xml_unlist, value)

un <- unlist(xml_df)
print(un)

# flat xml

# install.packages("flatXML")
# install.packages("devtools")
# require(devtools)
# install_github("https://github.com/jsugarelli/flatxml/")
# require(flatxml)
install.packages("xmlconvert", dependencies = TRUE)
require(xmlconvert)

fx <- fxml_importXMLFlat()
fxdf <- xml_to_df("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml")

```

## D E M Z

```{r PACKAGES}

require(xml2)

```

```{r PISKOVISTE}

xml1 <- as_list(read_xml("C:/Users/krizova/Documents/R/data/_piskoviste/10918171.xml"))
xml2 <- as_list(read_xml("P:/w4dz/migrace/boj/26272175.xml"))
xml3 <- as_list(read_xml("P:/w4dz/migrace/boj/26280130.xml"))
xml4 <- as_list(read_xml("P:/w4dz/migrace/boj/26282696.xml"))
xml5 <- as_list(read_xml("P:/w4dz/migrace/boj/26332340.xml"))

path_xml <- "P:/w4dz/migrace/"
fold <- "boj"
# fold <- "zmt"
Lx <- c("10918171.xml", "26272175.xml", "26280130.xml", "26282696.xml", "26332340.xml")

infoex <- data.frame()

for (i in Lx) {
  a <- as_list(read_xml(paste0(path_xml, fold, "/", i)))
  b <- tibble::as_tibble(a) %>%
  unnest_longer(import)
  c <- b %>%
  dplyr::filter(import_id == "PrirCisl") %>%
  unnest_wider(import, names_sep = "_")
  s <- c[1]
  infoex <- rbind(infoex, s)
}

```

*!!! NIZE JEN OVERENE A FUNKCNI PRIKAZY !!!*

```{r LOAD AND EDIT XML}

# L I S T   O F   F I L E S ^

Lxml_boj <- list.files("P:/w4dz/migrace/boj/", pattern = "xml") # 182
Lxml_zmt <- list.files("P:/w4dz/migrace/zmt/", pattern = "xml") # 59795

Lxml <- c(Lxml_boj, Lxml_zmt) # 59977

# L O O P   F O R   E X T R A C T I N G   S P E C.   I N F O

infoex <- data.frame() # info export dataframe

for (i in Lxml_boj) {
  a <- as_list(read_xml(paste0(path_xml, fold, "/", i))) # read xml
  b <- tibble::as_tibble(a) %>%  
  unnest_longer(import)
  c <- b %>%
  dplyr::filter(import_id == "Signatura") %>%
  unnest_wider(import, names_sep = "_")
  s <- c[1]
  infoex <- rbind(infoex, s)
}

```