### 2023-05-19 _notes_.Rmd
MUZBE ohrozeni

---
title: "RMD for notes"
subtitle: ""
author: "Katerina Krizova"
date:  "`r Sys.Date()`"
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

\newpage

# R general tips

- pojmenovavej df s prefixem, at se hezky radi v environmentu

# MIG tips

## naming conventions

- Mname = name of working column 
- MNAME = final name of column

## process

- *wip_tab* = work-in-progress tab (try everything out here)
- *mod_tab* = result of modifications, source for building od a mus_tab
- *mus_tab* = final CSV for import