### 2023-05-31 _GEN_axi_Rhacks.Rmd

---
title: "AXIELL R HACKS"
subtitle: "Uzitecne prikazy a funkce pro migrace dat"
author: "Kateřina Křížová"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---

```{r include = F}
knitr::opts_chunk$set(echo = F, eval = F)
```

\newpage

# *GENERAL*

## I M P O R T / E X P O R T

### Read txt

read.table

```{r READ TABLE}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.txt", header = TRUE, sep = ";")
```

### Read csv

read.csv : sep is always ","

```{r READ CSV}
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

### Read excel

```{r READ EXCEL}

library("readxl")
df <- read_excel("D:/VURV/R_hacks/data/simple_table_4read.xlsx" , sheet = 2, header = TRUE)
```

ISSUE: 'Warning: Expecting numeric in E1251 / R1251C5: got '41 (cm)'

SOLUTION: df <- read_excel(paste0(path_snd, "muzbe_rozmery_ptaci.xlsx"), sheet = 1, **col_types="text"**)

### Write excel

```{r}
write.xlsx(x = ToothGrowth,                       
           file = "ToothGrowth.xlsx")
```

### Write excel + sheets

```{r}
exi <- paste0(path_ces, proj, "/", proj, "_KAT.xlsx")

if(file.exists(exi)){
  print("File already exists !!!")
} else {
  f2s = exi
  write.xlsx(as.data.frame(SAVE1), file= f2s, sheetName="sparovano", row.names=FALSE)
  write.xlsx(as.data.frame(SAVE2), file= f2s, sheetName="duplicity", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE3), file= f2s, sheetName="nesparovanoMUS", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE4), file= f2s, sheetName="nesparovanoCES", append=TRUE, row.names=FALSE)
  write.xlsx(as.data.frame(SAVE0), file= f2s, sheetName="puv. vypis CES", append=TRUE, row.names=FALSE)
  rm(exi)
  rm(f2s)
  }
```

## N U M E R I C S

### Leading zeros -add

```{r LEADING ZEROS}

# package
require(stringr) # ::str_pad

# command
mutate(MPORC = str_pad(PorC_S, 6, pad = "0"))
```

mutate = create a new column  
MPORC = name of a new column
str_pad = [function] (https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_pad)
PorC_S = name of a source column (contains value '6' which we want add lead.zeros to)
6 = we want a final value to have exactly 6 places  
pad = "0" = empty places fill with "0"

### Leading zeros -remove

```{r}

mutate(Mcislo = sub("^0+", "", Mcislo))
```

### ID and sequence along

```{r PORADI}

v <- c("raz", "dva", "tři", "čtyři", "pět") 
v <- as.data.frame(v) %>% 
  mutate(por = seq_along(v),
         porplus = seq_along(v)+1)
```

### Round

round(x, #) = zaokrouhleni na # pocet decimalnich cisel
ceiling(x) = rounds values up to nearest integer
floor(x) = rounds values down to nearest integer
trunc(x) = truncates (cuts off) decimal places from values

```{r}

```

\newpage

## S T R I N G S

### Length

```{r STRING LENGTH}

mutate(Mpskp_len = nchar(pskup)) 
```

### Similarity

install.packages("RecordLinkage")
require("RecordLinkage")

levenshteinSim = funkce
autor_MUS, autor_AOPK = sloupce, ktere chci porovnat

```{r STRING SIMILARITY}

df %>% mutate(sim = levenshteinSim(autor_MUS, autor_AOPK))
```

### Detect case INsensitive

```{r STRING CASE SENS}

str_detect(MFN, regex("^FM\\s\\d+\\.jpg$", ignore_case = T))
```

### Separate

```{r SPLIT STRING}

# # Split name column into firstname and last name
# df[c('First Name', 'Last Name')] <- str_split_fixed(df$Name, ' ', 2)
#  
# # Rearrange columns and remove original name column
# df <- df[c('First Name', 'Last Name', 'State')]

separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ")

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge") # at first whitespace

separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge", remove = F) # ponecha puvodni sloupec
```

### Keep text before

```{r}
  mutate(blMCELED = sub("\\/.*", "", fiMTREE)) # ponechej text do prvni /
```

### Extract word

```{r}
mutate(boMROD = word(druh, 1)) # vezmi 1. slovo ze sloupce druh
```

### Remove parentheses

```{r}

mutate(celed_LAT = gsub("[()]", "", celed_LAT))
```

### Remove whitespace

```{r RMV WHITE SPACE}

# package
require(stringr) # ::str_pad

# command
mutate(MMATERIAL = str_trim(MMAT_AGG)) 
```

str_trim = [function](https://stringr.tidyverse.org/reference/str_trim.html)

### Unite

Creates a string of multiple column values with NAs.

```{r}

unite(sys_tree_BIOL, c("kmen", "trida", "rad", "celed", "rod", "druh_Nomen"),
      sep = "/", na.rm = T, remove = FALSE) 
```

unite = [function](https://www.rdocumentation.org/packages/tidyr/versions/0.8.2/topics/unite)  
sys_tree_BIOL = name of a new column  
c(***) = list of columns  
sep = sign for separation  
na.rm = skip columns containing NA  
remove = keep or delete source columns

\newpage

## D A T A F R A M E S

### g e n e r a l

#### Replace "" with NA

s pipe jsem zatim cestu nenasla

```{r REPLACE "" WITH NA}

taxon[taxon == ""] <- NA
```

#### Replace NA with ""

```{r REPLACE NA WITH ""}

ou <- taxon %>%
  replace(is.na(.), "")

ou <- taxon %>%
  mutate_all(~na_if(., ''))
```

#### str_trim pro cely df

```{r STR_TRIM DF}

df %>% 
  mutate(across(where(is.character), str_trim))
```

### c o l u m n s

#### Remove columns acc to pattern

```{r}

bot <- bot %>% 
  select(-contains("_sort"), -nvg_id)
```

#### Remove last n columns

```{r}

select(-mus_invc, -(last_col(offset = 34):last_col()))
```

#### Add prefix to column names

```{r}
colnames(kmen) <- paste0("kmen_", colnames(kmen))
```

#### Merge two columns with not complete information

column 1 has some NAs  
column 2 has also some NAs but in different rows than column 1
TASK: create one column with merged info

```{r}
#package
require(dplyr) # part of tidyverse

# command
mutate(kmen = coalesce(kmen_Nomen, trida_NadNomen))
```

coalesce = [function](https://www.rdocumentation.org/packages/dplyr/versions/1.0.10/topics/coalesce)

#### Relocate columns

```{r}
relocate(MINVC, .before = JineC_S) 
relocate(MINVCDO, .after = MINVC)
```

#### Shift columns

```{r}
rows_to_change <- a$zmena == "N" # filter rows to be shifted
a[rows_to_change,3:12] <- a[rows_to_change,2:12] # 3:12 nova lokace, 2:12 stara lokace
a[rows_to_change,2] <- NA # puvodni lukaci (diru) nahradit NA
```

### r o w s

#### DF top 10 rows

```{r}

mus_tab_cvic <- mus_tab %>% 
  slice(1:10)
```

#### Random sample of n

```{r}

ch <- sample_n(df, 10) # n=10
```

#### Drop NA rows in spec col

```{r}

sez_int <- sez %>% 
  drop_na(MINVChorni)  
```

#### Copy rows / multiply

1:nrow(df) = vsechny radky
, 1), ]) = jednou (tzn. ve vysledku dostaneme kazdy radek dvakrat)

```{r}

df <- rbind(df, df[rep(1:nrow(df), 1), ])
```

#### Duplicates

```{r}

#  https://community.rstudio.com/t/filter-out-all-rows-with-duplicate-values/41043/4

# Return names which have only a single row of data
data %>% 
  group_by(name) %>% 
  filter(n()==1)

# Return names which have more than one row of data
data %>% 
  group_by(name) %>% 
  filter(n() != 1)
```

\newpage

## D A T A B A S E S

### Connect db and read table

```{r}
# package 
require(RODBC)

# general

dbname <- paste0("---cesta+jmeno-databaze.accdb") 
con <- RODBC::odbcConnectAccess2007(dbname)      
RODBC::sqlTables(con)
tab <- RODBC::sqlFetch(con, "---jmeno-tabulky---")

# old func example
dbname <- paste0(path_mdb, "botanika_test.mdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

RODBC::sqlTables(con)                          # list tables in database

kmen <- RODBC::sqlFetch(con, "data") # read specified table

path_mdb <- "M:/03 klienti/kraj moravskoslezsky/muzeum beskyd - MUZBE/02 muzBE konverze/zdrojdat/BIO/MB_PVO_Botanika/"

```

*.mdb = old MS ACCESS  
*.accdb = new MS ACCESS

### List files from one or more folders

```{r}

# one
vypis <- as.data.frame(list.files("D:/MUZBE/Negativy/Image", full.names = T))

# more
vypis <- as.data.frame(list.files(c("D:/MUZBE/Obecné sbírky-4složky obrázků/Image_01",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_02",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_03",
                                    "D:/MUZBE/Obecné sbírky-4složky obrázků/Image_04"), 
                                  pattern = "jpg$|JPG$", full.names = T))

colnames(vypis)[1] <- "path"
```

## J O I N S

*inner_join* only keeps observations from x that have a matching key in y
*left_join()* keeps all observations in x.
*right_join()* keeps all observations in y.
*full_join()* keeps all observations in x and y.

[description and parameters](https://dplyr.tidyverse.org/reference/mutate-joins.html)

### Join only selected columns

```{r JOIN SEL COL}

left_join(biolib %>% dplyr::select('4nazevKategorie', '5nomen', '8nadrizenyTaxon'),
            by = c("MNOMEN" = "5nomen"), keep = T)
```

\newpage

# *REAL TASKS*

## Save mod_tab a mus_tab

imp = importni rozhrani 
f2s = file to save

```{r SAVE CSV AND OPER FILE}

# finalni csv

f2s <- paste0(path_csv, projekt, "_import_", imp, ".csv")

if(file.exists(f2s)){
  print("You've been here before!")
  rm(f2s)
} else {
  write.table(mus_tab, file = f2s,
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
  print("Import csv  W R I T T E N  !")
  rm(f2s)
  }

# upravena tabulka

mod2s <- paste0(path_csv, projekt, "_modif_", imp, ".csv")

if(file.exists(mod2s)){
  print("You've been here before!")
  rm(mod2s)
} else {
  write.table(mod_tab, file = mod2s, 
            quote = T, row.names = F, 
            sep = ";", dec = ",",              # desetinny oddelovac musi byt carka
            na = "", fileEncoding="cp1250")
  print("Operational file  S A V E D  !")
  rm(mod2s)
  }
```

## Uprava dataci

```{r DERIVE VZNIKPTU FROM SBIRKY, warning = F, eval = F}

dopr_tab <- sbirky %>% 
  select(MINVC, Autor_S, Role_S, Datace_S, MistoVz_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate(MDATACE = Datace_S) %>% 
  # yyyy
  mutate(MDAT_Y = ifelse(str_detect(MDATACE, "_00_00"), 
                         str_replace_all(MDATACE, "_00|-00", ""), NA)) %>% 
  # mm.yyyy
  # mutate(MDAT_YM = ifelse(str_detect(MDATACE,
  #                                    "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
  #                                   # format(anydate(substring(MDATACE, 1, 7), "%m.%Y")), ""))
  #                                   # substring(format(anydate(MDATACE), "%m.%Y"), 1, 7), ""))
  #                                   format(anydate(MDATACE), "%m.%Y"), ""))
  # mutate(MDAT_YM = format(anydate(MDATACE), "%m.%Y")) %>%
  mutate(MDAT_YM = substring(ifelse(str_detect(MDATACE,
                                             "^([0-9]{4})[_]([0-9]{1})([1-9]{1})[_]00$"),
                                  MDATACE, ""), 1, 7)) %>%
  mutate(MDAT_YM = format(anydate(MDAT_YM), "%m.%Y")) %>%
  # dd.mm.yyyy
  mutate(MDAT_YMD = ifelse(str_detect(MDATACE,
                          "^([0-9]{4})[_]([0-9]{2})[_]([0-9]{2})$"), 
                           format(anydate(MDATACE), "%d.%m.%Y"), NA)) %>% 
  # yy leta ss stoleti
  mutate(MDAT_LETA = ifelse(str_detect(MDATACE, "^([0-9]{4})[_]XX[_]([0-9]{2})$"),
                                    paste0(substring(dopr_tab$MDATACE, 9, 9), "0. léta ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # poloviny
  mutate(MDAT_2 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d2[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". polovina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # tretiny - NEVYSKYTUJE SE
  mutate(MDAT_3 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>%
  # poloviny - NEVYSKYTUJE SE
  mutate(MDAT_4 = ifelse(str_detect(MDATACE, "^00\\d\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". čtvrtina ",
                                           substring(dopr_tab$MDATACE, 3, 4), ". století"), NA)) %>% 
  # tretina roku
  mutate(MDAT_3r = ifelse(str_detect(MDATACE, "^\\d([1-9]{2})\\d[_]\\d3[_]XX"),
                                    paste0(substring(MDATACE, 6, 6), ". třetina roku ",
                                           substring(dopr_tab$MDATACE, 1, 4)), NA)) %>% 
  # DATACE FINAL
  mutate(MDATACE_FIN = coalesce(MDAT_Y, MDAT_YM, MDAT_YMD, MDAT_LETA, MDAT_2, MDAT_3r)) %>% 
  mutate(MDATACE_FIN = coalesce(MDATACE_FIN, MDATACE))
```

## Tvorba inventarniho cisla

```{r MINVC}

sbirky <- read_excel(paste0(path_data, "/Sbirky.xlsx")) %>%

  select(-MSUBOD, -MSUBDO, -MINVC, -MINVCDO) %>%    # get rid of .accdb columns

  mutate(MPORC = str_pad(PorC_S, 6, pad = "0"),
         MPORCDO = str_pad(HPorC_S, 6, pad = "0"),
         MSUB = ifelse(Lomeni_S == "_", NA, Lomeni_S),
         MSUBDO = HLomeni_S, 
         MINVC = case_when(is.na(MSUB) ~ paste0(Rada_S, "-", MPORC),
                             TRUE ~ paste0(Rada_S, "-", MPORC, "/", MSUB)),
         MINVCDO = case_when(is.na(MSUBDO) ~ "",
                             TRUE ~ paste0(Rada_S, "-", MPORCDO, "/", MSUBDO))) %>% 
  # select(-MPORC, -MPORCDO) %>%
  relocate(MINVC, .before = JineC_S) %>% 
  relocate(MINVCDO, .after = MINVC)

# one step

MINVC = paste0("M", str_pad(invc, 5, pad = "0"))

```

## Taxon -> rod, druh, autor

```{r SPLIT TAXON}

separate(taxon, into=c("Mrod", "Mdruh", "Mautor") , sep=" ", extra = "merge", remove = F)
```

## Check import KD

```{r KD IMPORT CHECK}

mus_imp <- read_excel(paste0(path_kd, v, "/KontextovyDokument_export (1).xlsx")) %>%  
  mutate(MFN = sub("_original", "", URL))

jpg_cs <- jpg %>% mutate(MFN = ifelse(str_detect(MFN, "JPG"), sub("JPG", "jpg", MFN), MFN))

imp_check <- left_join(jpg_cs, mus_imp, by = c("MFN" = "MFN"), keep = T)

na <- imp_check %>% filter(is.na(Číslo))

```

\newpage

# *NEED TO COVER*

dopr_tab <- sbirky %>% 
  select(MINVC, Technika_S) %>% 
  mutate(MEMPTY = "") %>% 
  mutate('2sbirkaCisloEvidInt' = sbirka) %>% 
  mutate('3podsbirkaCislo' = podsbirka) %>% 
  mutate(MTECH = Technika_S) %>% 
  separate(MTECH,into=c("MTECH1", "MTECH2", "MTECH3", "MTECH4", "MTECH5"),
           sep=",(?! tuší)") %>% # separator is ',' BUT ONLY IF NOT FOLLOWED BY ' tuší' -> solves a single case of 'kresba tužkou, tuší a bělobou' which should remain as one entry
  pivot_longer(
    cols = MTECH1:MTECH5, 
    names_to = "MTECH_POR", # poradi techniky
    values_to = "MTECH_AGG",
    values_drop_na = TRUE) %>% 
  mutate(MTECHNIKA = str_trim(MTECH_AGG)) %>% # remove white space
  select(-MTECH_AGG) %>% 
  mutate(MTECH_POR = substr(MTECH_POR, 6, 6))

//

plant <- db_tab %>%

   separate(celed,into=c("celed_LAT", "celed_CZ") , sep=" ") %>% 
   separate(druh,into=c("rod_LAT", "druh_LAT", "rest"), sep = "\\s(?!x )", 
            extra = "merge", remove = FALSE) %>% 
   mutate(druh_LAT = paste(rod_LAT, druh_LAT),
          celed_CZ = gsub("[()]", "", celed_CZ),
          autor = ifelse(str_detect(rest, "^[[:upper:]]") | str_detect(rest, "^\\W"), rest,""),
          druh_LAT = gsub("x$", "", druh_LAT))

//

 mutate_all( ~ str_replace_all(., pattern = ";", replacement = " ")) %>% # removes semicolons
  mutate_all( ~ str_replace_all(., pattern = "[\r\n]", replacement = " ")) #removes paragraphs