### 2024-07-02 NM_analyza.Rmd
Hacker attack commit

---
title: "AXIELL piskoviste"
subtitle: "Prostredi pro zkouseni novych postupu"
author: "Kateřina Křížová"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

---
# B A S E

```{r}
require(tidyverse)
require(readxl)
require(stringr) # ::str_split_fixed
require(stringi) # ::stri_detect_fixed - detect empty string
require(mdbr) # ::read_mdb
require(RODBC) # ::read_mdb
require(odbc) # ::read_mdb
require(xlsx) # write excel in sheets

path_nm <- "M:/03 klienti/narodni muzeum - NM/2024_analyzaNM/KK_NM_workspace/_workspace/"
```

# MARK 21

```{r}

install.packages("devtools")
devtools::install_github("dojobo/marc21r")
require(marc21r)

```

```{r}
# Specify the path to your MARC 21 file
marc_file <- paste0(path_nm, "DATAzSHP/KNM1/export-KNM-Dlouhodobevypujcky-240523-1067zaznamu.txt")

# Read the MARC 21 data
marc_records <- marc21r::read_marcxml(marc_file)

# Function to extract a specific field from a MARC record
extract_field <- function(record, tag) {
  field <- record$fields[[tag]]
  if (is.null(field)) return(NA)
  if (is.list(field)) return(paste(sapply(field, function(x) x$value), collapse = " "))
  return(field$value)
}

# Extract specific fields from the MARC data
marc_df <- data.frame(
  Title = sapply(marc_records, extract_field, "245"),
  Author = sapply(marc_records, extract_field, "100"),
  PublicationInfo = sapply(marc_records, extract_field, "260"),
  stringsAsFactors = FALSE
)

# Clean up the data frame
marc_df <- marc_df %>%
  mutate(
    Title = sub("^[^a-zA-Z0-9]*", "", Title),
    Author = sub("^[^a-zA-Z0-9]*", "", Author),
    PublicationInfo = sub("^[^a-zA-Z0-9]*", "", PublicationInfo)
  )

# View the resulting data frame
print(marc_df)

```