### 2025-02-03 TF_UAV.Rmd

---
title: "UAV / multispectral data processing"
author: "Katerina Krizova"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 3
---

\newpage

```{r SETUP PACKAGES INDEX, include=FALSE}

# RMARKDOWN SETUP

knitr::opts_chunk$set(echo = F, message=FALSE, fig.dim = c(8, 4))

# PACKAGES

# list.of.packages <- c("devtools","mapview", "leaflet","randomcoloR","Rcpp", "sf", "tidyverse", "whitebox")
# install.packages(list.of.packages)
# install.packages('raster', repos='https://rspatial.r-universe.dev')
# 
# lapply(list.of.packages, require, character.only = TRUE)
# require(raster)
# 
# rm(list.of.packages)

# !!!!!!!!!!!!!!!!!!!!!!!! S W I T C H !!!!!!!!!!!!!!!!!!!!!!!!

index <- "ndvi" #pse
# index <- "sipi" # BAND RESOLUTION ISSUE TO SOLVE
# index <- "ndre" #pse
# index <- "gndvi" #pse
# index <- "evi" # BAND RESOLUTION ISSUE TO SOLVE
# index <- "tgi" #pse
# index <- "vari" #pse
# index <- "gli" # BAND RESOLUTION ISSUE TO SOLVE
# index <- "exg"
# index <- "ngrdi" #pse
# index <- "cvi" #pse

# kat indexy
# index <- "exr"
# index <- "exgexr"
# index <- "savi"
# index <- "sr"
# index <- "mcari"

# !!!!!!!!!!!!!!!!!!!!!!!! S W I T C H !!!!!!!!!!!!!!!!!!!!!!!!

```

# PATHS AND PARAMS

*CRS*

EPSG 32633 (UTM 33N)

```{r PATH PARAMS}

# !!!!!!!!!!!!!!!!!!   S W I T C H   !!!!!!!!!!!!!!!!!!

# project <- "jecmen"
project <- "psenice"

# !!!!!!!!!!!!!!!!!!   S W I T C H   !!!!!!!!!!!!!!!!!!

# path <- "C:/Users/krizovak/Documents/TF_UAV/" # home folder

path_in <- paste0(path, "data/", project, "/")
path_alt <- paste0("D:/VURV/_____odchod/TF_UAV/data/", project, "/")
path_out <- paste0(path, "results/", project, "/")

if (project == "jecmen") {
    Lterm <- c(
      "220510", "220514", "220521", "220529", "220605", "220612",
               "220618", "220625", "220702", "220709", "220716"
      ) 
} else if (project == "psenice") {
    Lterm <- c(
      "220410", "220423", "220429", "220510", "220522", "220529",
    "220605", "220612",
    "220618", "220625", "220702", "220709", "220716"
    )  
} else {print("UNSUCCESSUL")}

```

# VECTOR LAYERS

```{r LOAD POINTS CREATE BUFFER}

# load points and convert to shp

# pts_csv <- read.csv(paste0(path_in, "pts_", project, ".csv"), sep = ";", 
#                     check.names = F, encoding = "Windows-1250") %>% 
#   na.omit()
# pts4326 <- sf::st_as_sf(pts_csv, coords = c("lon", "lat"))
# ggplot()+
#   geom_sf(data = pts4326)+
#   theme_classic()

# pekar 2025_1
pts_xlsx <- read_excel(paste0(path, "body.xlsx"), sheet = "body") %>% 
  na.omit() %>% 
  mutate(bod_id = paste0(body_a, "_", body_b))

# pts4326 <- sf::st_as_sf(pts_xlsx, coords = c("lon", "lat"))
pts4326 <- sf::st_as_sf(pts_xlsx, coords = c("x", "y"))
g <- ggplot()+
  geom_sf(data = pts4326)+
  theme_classic()
g
print(pts4326)

pts_csv <- read.csv(paste0(path_in, "pts_", project, ".csv"), sep = ";", 
                    check.names = F, encoding = "Windows-1250") %>% 
  na.omit()
pts4326 <- sf::st_as_sf(pts_csv, coords = c("lon", "lat"))
ggplot()+
  geom_sf(data = pts4326)+
  theme_classic()

# transform CRS to 32633

sf::st_crs(pts4326) = 4326
pts <- sf::st_transform(pts4326, crs = 32633)

# library(ggsflabel)

g <- ggplot()+
  # geom_sf(data = buf, color = "gold", fill = "gold", size = 2)+
  geom_sf(data = pts_xlsx, size = 2)+
  # geom_text(data = pts_xlsx, aes(lon, lat, label = ID))+
  # geom_text(data = pts_xlsx, aes(lon, lat, label = bod_id))+
  coord_sf(crs = 32633)+
  labs(title = paste0("Sampling points"))+
  theme_classic()+
  theme(plot.title=element_text(hjust=0.5))
g
ggsave(g, filename = paste0(path_out, "pts_map.jpg"))

# remove unnecessary objects

rm(pts_csv)
rm(pts4326)

```

# RASTER LAYERS

```{r LOAD BANDS, eval = F}

rgb <- raster(paste0(path_in, "220529_rgb.tif"), quiet = T) # 4 bands: b1:B b2:G b3:R b4:EMPTY
ms <- raster(paste0(path_in, "220529_ms.tif"), quiet = T) # 5 bands: b1:G b2:R b3:NIR b4:REDEDGE b5:EMPTY

ms <- raster(paste0(path_in, "220410/"), pattern = "*_M*.tiff"), quiet = T) # 5 bands: b1:G b2:R b3:NIR b4:REDEDGE b5:EMPTY

# RGB 
# blue <- raster(paste0(path_in, "220529_rgb.tif"), band = 1, quiet = T)
# green <- raster(paste0(path_in, "220529_rgb.tif"), band = 2, quiet = T)
# red <- raster(paste0(path_in, "220529_rgb.tif"), band = 3, quiet = T)
# rgb_empty <- raster(paste0(path_in, "220529_rgb.tif"), band = 4, quiet = T)

# MS

green <- raster(paste0(path_in, "220529_ms.tif"), band = 1, quiet = T)
red <- raster(paste0(path_in, "220529_ms.tif"), band = 2, quiet = T)
re <- raster(paste0(path_in, "220529_ms.tif"), band = 3, quiet = T)
nir <- raster(paste0(path_in, "220529_ms.tif"), band = 4, quiet = T)
ms_empty <- raster(paste0(path_in, "220529_ms.tif"), band = 5, quiet = T)

# pekar 2025_1

green <- raster(paste0(path, "ndvi_20240415.tif"), band = 1, quiet = T)
red <- raster(paste0(path, "ndvi_20240415.tif"), band = 2, quiet = T)
re <- raster(paste0(path, "ndvi_20240415.tif"), band = 3, quiet = T)
nir <- raster(paste0(path, "ndvi_20240415.tif"), band = 4, quiet = T)
ms_empty <- raster(paste0(path, "ndvi_20240415.tif"), band = 5, quiet = T)

ndvi <- raster(paste0(path, "ndvi_20240415.tif"), quiet = T)
nlayers(ndvi)
```

# CALCULATE VI

```{r CALCULATE VI LOOP, warning = F}

# imagery

# 4 bands: b1:B b2:G b3:R b4:EMPTY
# 5 bands: b1:G b2:R b3:NIR b4:REDEDGE b5:EMPTY

# index calculation LOOP

for(i in Lterm){
  blue <- raster(paste0(path_in, i, "/rgb.tif"), band = 1, quiet = T)
  green <- raster(paste0(path_in, i, "/rgb.tif"), band = 2, quiet = T)
  c_red <- raster(paste0(path_in, i, "/rgb.tif"), band = 3, quiet = T)
  m_green <- raster(paste0(path_in, i, "/ms.tif"), band = 1, quiet = T)
  red <- raster(paste0(path_in, i, "/ms.tif"), band = 2, quiet = T)
  edge <- raster(paste0(path_in, i, "/ms.tif"), band = 3, quiet = T)
  nir <- raster(paste0(path_in, i, "/ms.tif"), band = 4, quiet = T)
  new1 <- raster(paste0(path_in, i, "/020722.data.tif"), band = 1, quiet = T) # blue ?
  new2 <- raster(paste0(path_in, i, "/020722.rgb.tif"), band = 1, quiet = T)
# # alternative source
  # blue <- raster(paste0(path_alt, i, "/rgb.tif"), band = 1, quiet = T)
  # green <- raster(paste0(path_alt, i, "/rgb.tif"), band = 2, quiet = T)
  # c_red <- raster(paste0(path_alt, i, "/rgb.tif"), band = 3, quiet = T)
  # m_green <- raster(paste0(path_alt, i, "/ms.tif"), band = 1, quiet = T)
  # red <- raster(paste0(path_alt, i, "/ms.tif"), band = 2, quiet = T)
  # edge <- raster(paste0(path_alt, i, "/ms.tif"), band = 3, quiet = T)
  # nir <- raster(paste0(path_alt, i, "/ms.tif"), band = 4, quiet = T)
#index
  if(index == "evi") {
    vi <- 2.5*(nir-red)/(nir + 6*red - 7.5*blue + 1)
    } else if(index == "cvi") {
      vi <- (nir/m_green)*(red/m_green)
    } else if(index == "exg") {
      vi <- 2*(green/(green+blue+c_red))-(c_red/(green+blue+c_red))-(blue/(green+blue+c_red))
    } else if(index == "exr") {
      vi <- 1.4*((green/(green+blue+red))-(blue/(green+blue+red)))
    } else if(index == "exgexr") {
      vi <- exg-exr
    } else if(index == "gli") {
      vi <- (2*green-red-blue)/(2*green+red+blue)
    } else if(index == "gndvi") {
      vi <- (nir-m_green)/(nir+m_green)
    } else if(index == "mcari") {
      vi <- ((nir-red)-0.2*(nir-green))*(nir/red)
    } else if(index == "ndre") {
      vi <- (nir - edge)/(nir + edge)
    } else if(index == "ndvi") {
      vi <- (nir-red)/(nir+red)
    } else if(index == "ngrdi") {
      vi <- (m_green-red)/(m_green+red)
    } else if(index == "savi") {
      vi <- (1.5*(nir-red))/(nir+red+0.5)
    } else if(index == "sipi") {
      vi <- (nir-blue)/(nir-red)
    } else if(index == "sr") {
      vi <- nir/red
    } else if(index == "tgi") {
      vi <- green-0.39*c_red-0.61*blue
    } else if(index == "vari") {
      vi <- (green-c_red)/(green+c_red-blue)
    }else {(print("UNSUCCESSFUL"))}
  dir.create(paste0(path_out, index))
  # writeRaster(vi, paste0(path_out, index, "/", i, "_", index, ".tif"), format="GTiff", overwrite=TRUE)
  # writeRaster(vi, paste0("D:/VURV/_____odchod/", i, "_", index, ".tif"), format="GTiff", overwrite=TRUE)
}

```

# ZONAL STATISTICS

```{r ZONAL STAT, eval = F}

# !!!!!!!!!!!!!!!!!!   S W I T C H   !!!!!!!!!!!!!!!!!!

# distance <- 1 # psenice # jecmen
# distance <- 5 # psenice
distance <- 10 # psenice

# !!!!!!!!!!!!!!!!!!   S W I T C H   !!!!!!!!!!!!!!!!!!

# create buffer

buf <- sf::st_buffer(pts, dist = distance) 
# buf <- sf::st_transform(pts, crs = 4326)

# zonal stat folder and file

VI_results <- data.frame() # create empty dataframe for results

# zonal stat LOOP

cbind.fill <- function(...){
    nm <- list(...) 
    nm <- lapply(nm, as.matrix)
    n <- max(sapply(nm, nrow)) 
    do.call(cbind, lapply(nm, function (x) 
        rbind(x, matrix(, n-nrow(x), ncol(x))))) 
} # FUNCTION from: https://stackoverflow.com/questions/7962267/cbind-a-dataframe-with-an-empty-dataframe-cbind-fill

for(i in Lterm){
  vi <- raster(paste0(path_out, index, "/", i, "_", index, ".tif"), quiet = T)
  vi_mean <- raster::extract(vi, buf, fun='mean', na.rm=TRUE, df=TRUE, weights = TRUE)
  df <- data.frame(vi_mean) 
  colnames(df)[2] <- i
  VI_results <- cbind.fill(VI_results, df)
}

# zonal stat tab

VI_results <- unique(VI_results, MARGIN=2)
VI_results <- as.data.frame(VI_results)
VI_results$ID <- as.factor(VI_results$ID)
# buf$Bod <- as.factor(buf$Bod)
buf$ID <- as.factor(buf$ID)
# levels(VI_results$ID) = levels(buf$Bod)
levels(VI_results$ID) = levels(buf$ID)

write.table(VI_results, paste0(path_out, index, "_mean_buffer_", distance, ".txt"), row.names=F, quote = F)

```

# ZONAL STATISTICS LOOP

```{r ZONAL STAT LOOP}

cbind.fill <- function(...){
    nm <- list(...) 
    nm <- lapply(nm, as.matrix)
    n <- max(sapply(nm, nrow)) 
    do.call(cbind, lapply(nm, function (x) 
        # rbind(x, matrix(, n-nrow(x), ncol(x))))) 
        rbind(x, matrix(n-nrow(x), ncol(x))))) 
} # FUNCTION from: https://stackoverflow.com/questions/7962267/cbind-a-dataframe-with-an-empty-dataframe-cbind-fill

if(project == "jecmen"){
  Ldist <- c(1)
} else if (project == "psenice"){
  Ldist <- c(1,5,10)
} else {print("UNSUCCESSUL")}
cust_seq <-1:12

for(i in Ldist){
 buf <- sf::st_buffer(pts, dist = i) 
 VI_results <- data.frame(cust_seq) # create empty dataframe for results
 for(j in Lterm){
  vi <- raster(paste0(path_out, index, "/", j, "_", index, ".tif"), quiet = T)
  vi_mean <- raster::extract(vi, buf, fun='mean', na.rm=TRUE, df=TRUE, weights = TRUE)
  df <- data.frame(vi_mean) 
  colnames(df)[2] <- j
  VI_results <- cbind(VI_results, df)
  VI_results <- unique(VI_results, MARGIN=2)
  VI_results <- as.data.frame(VI_results)
  # VI_results$ID <- as.factor(VI_results$ID)
  # # buf$Bod <- as.factor(buf$Bod)
  # buf$ID <- as.factor(buf$ID)
  # # levels(VI_results$ID) = levels(buf$Bod)
  # levels(VI_results$ID) = levels(buf$ID)
  write.table(VI_results, paste0(path_out, index, "_mean_buffer_", i, ".txt"), row.names=F, quote = F)
  }
}

```