### 2025-02-03 TF_UAV_pekar.Rmd

---
title: "Pekar 2025/1"
output: pdf_document
date: "2025-01-30"
---

# packages

```{r}

# install.packages("devtools")
# install.packages("usethis")
# install.packages("mapview")
# install.packages("leaflet")
# install.packages("randomcoloR")
# install.packages("Rcpp")
# install.packages("sf")
# install.packages("whitebox")
# install.packages("readxl")
# install.packages("raster")

require(tidyverse)
require(devtools)
require(usethis)
require(mapview)
require(leaflet)
require(randomcoloR)
require(Rcpp)
require(sf)
require(whitebox)
require(readxl)
require(raster)
```

# paths and params

```{r}
path <- "D:/pekar/" # home folder 
```

# pts

```{r}

pts_xlsx <- read_excel(paste0(path, "body.xlsx"), sheet = "body") %>% 
  na.omit() %>% 
  select(-1) %>% 
  mutate(bod_id = paste0(body_a, "_", body_b))

# pts4326 <- sf::st_as_sf(pts_xlsx, coords = c("lon", "lat"))
pts4326 <- sf::st_as_sf(pts_xlsx, coords = c("x", "y"))
g <- ggplot()+
  geom_sf(data = pts4326)+
  theme_classic()
g

```

# visible

```{r}
rgb <- raster(paste0(path, "visible_20240415.tif"), quiet = T) # 4 bands: b1:B b2:G b3:R b4:EMPTY
nlayers(rgb)

# ms <- raster(paste0(path_in, "220529_ms.tif"), quiet = T) # 5 bands: b1:G b2:R b3:NIR b4:REDEDGE b5:EMPTY
```
# ndvi

```{r}
ndvi <- raster(paste0(path, "ndvi_20240415.tif"), quiet = T)
nlayers(ndvi)
```